---
title: "Basics"
format: html
editor_options: 
  chunk_output_type: console
---

## Introduction to RStudio IDE
- Source
- Console
- Environment
- History
- Files
- Plots
- Help

## How to run R code
- from the console
- from the source code editor
- from the command line

```{r}
cat('Hello, world!')

x <- 'Zhang San'
cat('Hello, ', x, '!', sep='')
```

```{bash}
Rscript --vanilla tmp/hello.R "Li Si"
```

## Install packages
Within R
```{r}
install.packages('package_name')
```

With miniforge (suitable on servers or for packages with complex dependency)
```{bash}
mamba install conda-forge::r-tidyr
mamba install bioconda::bioconductor-deseq2
```

load installed packages
```{r}
library(tidyverse)
```

## Getting help
```{r}
# option 1
# get internal documentation
?pluck
help(pluck)

# option 2
# see source code
help
pluck

# option 3
# seach online (google or other websites)
# preferred sources: stackoverflow, github

# option 4
# ask large language models

# option 5
# ask an experienced user for help
# be sure to prepare a reprex (minimal reproducible example)
```

Tips:
> Invest in yourself. You should also spend some time preparing yourself to solve problems before they occur. Investing a little time in learning R each day will pay off handsomely in the long run.

## Plots
```{r}
plot(1:10)

pdf('tmp/110.pdf', width = 4, height = 3)
plot(1:10)
dev.off()

library(ggplot2)

ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point() + 
  theme_minimal()
ggsave('tmp/mtcars_wt_mpg.png', width = 4, height = 3)
```

## projects
- how to create a new project
- it's better to keep all relevant files in the same project folder.

project organization
```
.
├── 01_noise_ploidy_simulation.ipynb
├── 02_scrna_data_integration.ipynb
├── README.md
├── S01_10xv3_raw_to_cellbarcode_matrix.qmd
├── S02_10xv3_cellbender_scdblfinder.qmd
├── S03_cbmatrix_qc.ipynb
├── data                                                   # raw data; downloaded data
│   ├── HAv3.1.57.gtf.txinfo.tsv
│   ├── LiuJiamin_GeneList_20231127.xlsx
│   ├── LiuJiamin_GeneList_20231127_PeakSummary.xlsx
│   ├── NEE_2023_Traniello_SI.xlsx
│   ├── apiscc_cau.txinfo.tsv
├── figures                                                # visualizations with reasonable names
│   ├── acer_all_cluster_deg_filtered.pdf
│   ├── acer_all_cluster_deg_good_markers.pdf
│   ├── acer_all_cluster_fano_pairs.png
│   ├── acer_all_cluster_marker_exp.pdf
│   ├── acer_before_integration.pdf
│   ├── acer_integration.pdf
│   └── seurat_sctransform_cca_integration.pdf
├── results                                                # processed data; analysis results;
│   ├── FCA_cell_type_markers_clean.tsv
│   └── orthofinder_ace_orthologs_pairwise_aceK.tsv.gz
├── src                                                    # scripts to be run from cmdline
│   ├── simulation.jl
│   └── run_scDblFinder.R
└── tmp                                                    # temporary files; can be deleted later
    ├── 01_simulation-Copy1.ipynb
    ├── README.md
    ├── Untitled.ipynb
    └── acer_all_clu_0_cnt_stats.tsv
```

## scripts
- suffix: pure R scirpts: `.R` or markdown (`.Rmd`, `.qmd` or `.md`)
- name: meanningful, lower_case, no space or special characters; preferrably numbered.
    - like: `01_prepare_data.qmd`, `02_xxx_analysis.qmd`, `03_yyy_analysis.qmd`

## Other tips
- save your open files frequently.
- backup your code (for example, github/gitlab/gitee/gitlink)
- write comments
- leave some space

## coding styles

### Assignment

- Only use `=` when setting function arguments;
- Never use `->`;

```{r}
# Good
x <- 5

# syntactically correct but Bad
x = 5
5 -> x
```


### names (of variables or functions)
```{r}
# prefer: simple, meaningful, lower_case;
iris_setosa <- iris |> filter(Species == 'setosa')
# avoid:
subset_of_setosa_in_iris_dataset <- iris |> filter(Species == 'setosa')
irisSetosa <- filter(iris, Species == 'setosa')
IrisSetosa <- filter(iris, Species == 'setosa')
iRISSETOSA <- filter(iris, Species == 'setosa')
a <- filter(iris, Species == 'setosa')
b <- filter(iris, Species == 'versicolor')
ab <- bind_rows(a, b)

# Bad: avoid using names of base functions or variables
T <- FALSE
c <- 10
mean <- function(x) sum(x)

```

### spaces
```{r}
# good
z <- (a + b)^2 / d
# bad
z<-( a + b ) ^ 2/d

# good
mean(x, na.rm = TRUE)
# bad
mean (x ,na.rm=TRUE)    # sapce after ",", not before ahead of ","
mean( x ,na.rm=TRUE)
mean( x, na.rm=TRUE )

# Good
x[, 1]
# Bad
x[,1]
x[ ,1]
x[ , 1]

# except these operators: ::, :::, $, @, ^, [, [[, unary - (like "-4"), unary + (like "+4");
# Most binoary operators should have spaces around them: ==, +, -, <- etc.
```

### Indentation

indent properly
```{r}
# good
if (y == 0) {
  if (x > 0) {
    log(x)
  } else {
    message("x is negative or zero")
  }
} else {
  y^x
}

# bad
if (y == 0)
{
    if (x > 0) {
      log(x)
    } else {
  message("x is negative or zero")
    }
} else { y ^ x }
```

`|>` or `%>%` should always have a space before it and should typically be the last thing on a line. After the first step of the pipeline, indent each line by two or four spaces.

(similar rules apply when composing figures with `+` using ggplot2)

```{r}
# good
iris |> 
  janitor::clean_names() |>
  group_by(species) |> 
  summarize(across(everything(), \(x) mean(x, na.rm = TRUE))) |> 
  arrange(sepal_width)

# bad
iris|>janitor::clean_names()|>group_by(species)|>summarize(across(everything(), \(x) mean(x, na.rm = TRUE)))|>arrange(sepal_width)
```

### Comments
Each line of a comment should begin with the comment symbol and a single space: `# `.

## Suggested readings
- more on coding styles: https://style.tidyverse.org/index.html

## References
Some examples or notes in this page were directly copied from the following sources:

- https://style.tidyverse.org/index.html by Hadley Wickham
- https://r4ds.hadley.nz/ by Hadley Wickham, Mine Çetinkaya-Rundel, and Garrett Grolemund.

